---
output:
    md_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Status

**Early development**
**Version**: 0.0.1

[![Travis](https://travis-ci.org/TheoreticalEcosystemEcology/alienR.svg?branch=master)](https://travis-ci.org/TheoreticalEcosystemEcology/alienR)
[![Build status](https://ci.appveyor.com/api/projects/status/sk3sbvusvcyy0at0?svg=true)](https://ci.appveyor.com/project/TheoreticalEcosystemEcology/alienR/build/1.0.7)
[![codecov](https://codecov.io/gh/TheoreticalEcosystemEcology/alienR/branch/master/graphs/badge.svg)](https://codecov.io/gh/TheoreticalEcosystemEcology/alienR)


## ALIEN: The aim of the package is to predict All Links In Ecological Networks

## Installation

To install the development version, use the **devtools** package. 

```r
if (!require(devtools))
devtools::install_github("TheoreticalEcosystemEcology/alienR")
```

## Incoming content

### Methods implemented

- Direct matching centrality (DG)
    - binomial
    - poisson
    - RF
- Indirect matching centrality (latent ~ traits) (DG)
- Niche binary (DG)
- Niche probabilistic (DG)
- iEat recommend (DB)
- iEat gap-filling   (DB)  
- JSDM (GB)
- Harris (KC)
- Markov logic (A.T.?)
- 4th corner (GB)
- Ben's magic (BW)
- Ives' phylo (GB)

**Deadline:** Package ready by Christmas

## Generate package metadata

All enhancements to the README has to be written in the Readme.Rmd file.
To build the metadata files (i.e. README, NAMESPACE, DESCRIPTION):

```r
## Install packages
pkgs = c('devtools','roxygen2','testthat','formatR','rmarkdown')
for(pkg in pkgs) if(!require(pkg)) install.packages(pkg)

## Make sure you're working directory is set up correctly
getwd()

## Create README file
rmarkdown::render("README.Rmd", "all", quiet=TRUE)

## Format the R code (e.g. indentation, assignment, etc.)
formatR::tidy_dir("./R",arrow = getOption("formatR.arrow",TRUE))
formatR::tidy_dir("./tests/test_that", arrow = getOption("formatR.arrow",TRUE))

## Load your current version of the package
devtools::load_all(".")

## Create man pages
devtools::document(".")

## Testing the code with testthat package
devtools::test()
```

### For developers on MacOS and Linux only

All the steps described above are embedded in R script and integrated in a makefile:

```bash
cd path/to/alienR
make
```


## Contribute to this package

```{r child = 'CONTRIBUTION'}
```


## License: MIT
