---
title: "Salix_Galler_Parasitoids"
author: "Guillaume Blanchet"
date: "14/05/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This vignette was made to be an example of how the package can be applied to real-world data. As such, we focused on the *Salix*-galler-parasitoid data published by Kopelke et al. (2017) for which additional information on phylogeny was added. The full data was used to perform the analysis in Wooton et al. (2021).

# Data

The data is part of the `alien` R package and has already been formatted.

```{r}
library(alien)

data(salixGal)
data(galPara)
```

# `fitKNN`

## Interaction data only
```{r, warning=FALSE}
# Salix-galler
salixGalKNNInter <- fitKNN(salixGal,
                           distFrom = "jaccard",
                           distTo = "jaccard",
                           nNeig = 3,
                           weight = 0)

logLik(salixGalKNNInter)
tjur(salixGalKNNInter) 

# Galler-parasitoid
galParaKNNInter <- fitKNN(galPara,
                          distFrom = "jaccard",
                          distTo = "jaccard",
                          nNeig = 3,
                          weight = 0)

logLik(galParaKNNInter)
tjur(galParaKNNInter)
```

## Interaction data with traits
```{r, warning=FALSE}
# Salix-galler
salixGalKNNInterTrait <- fitKNN(salixGal,
                                distFrom = "jaccard",
                                distTo = "jaccard",
                                distTraitFrom = "euclidean",
                                distTraitTo = "euclidean",
                                nNeig = 3,
                                weight = 1)

logLik(salixGalKNNInterTrait)
tjur(salixGalKNNInterTrait) 

# Galler-parasitoid
galParaKNNInterTrait <- fitKNN(galPara,
                               distFrom = "jaccard",
                               distTo = "jaccard",
                               distTraitFrom = "euclidean",
                               distTraitTo = "euclidean",
                               nNeig = 3,
                               weight = 1)

logLik(galParaKNNInterTrait)
tjur(galParaKNNInterTrait) 
```

# Needs to be redone
## Interaction data with phylogeny 
# Needs to be redone
```{r, warning=FALSE}
# Salix-galler
salixGalKNNInterPhylo <- fitKNN(salixGal,
                                distFrom = "jaccard",
                                distTo = "jaccard",
                                nNeig = 3,
                                weight = 1,
                                phylo=TRUE)

logLik(salixGalKNNInterPhylo)
tjur(salixGalKNNInterPhylo)

# Galler-parasitoid
galParaKNNInterPhylo <- fitKNN(galPara,
                               distFrom = "jaccard",
                               distTo = "jaccard",
                               nNeig = 3,
                               weight = 1,
                               phylo = TRUE)

logLik(galParaKNNInterPhylo)
tjur(galParaKNNInterPhylo)
```

# `fitRF`

## Traits
```{r}
##############
# Salix-galler
##############

# Base of formula
salTraits <- paste(colnames(salixGal$traitFrom), collapse = "+")
galTraits <- paste(colnames(salixGal$traitTo), collapse = "+")

# Formula
FormulaSalGal <- as.formula(paste("~", salTraits, "+", galTraits))

# Model
salixGalRFTrait <- fitRF(salixGal,
                         formula = FormulaSalGal,
                         ntree = 2000,
                         nodesize = 1)

# Loglikelihood and Tjur's D
logLik(salixGalRFTrait)
tjur(salixGalRFTrait)

###################
# Galler-parasitoid
###################
# Base of formula
galTraits <- paste(colnames(galPara$traitFrom), collapse = "+")
parTraits <- paste(colnames(galPara$traitTo), collapse = "+")

# Formula
FormulaGalPara <- as.formula(paste("~", galTraits, "+", parTraits))

# Model
galParaRFTrait <- fitRF(galPara,
                        formula = FormulaGalPara,
                        ntree=2000,
                        nodesize = 1)

# Loglikelihood and Tjur's D
logLik(galParaRFTrait)
tjur(galParaRFTrait)
```

## Phylo
This needs to be done after answers from Kate.

# `fitGLM`

```{r}
################
# Salix - Galler
################
# Construct squared traits for numerical traits
salixGalSq <- polyTrait(salixGal)

#==============
# Build formula
#==============
# Select traits
salixTr <- c("TREE.VOLUME",
             "TREE.VOLUME_Sq",
             "LEAF.THICKNESS",
             "GLUCOSIDES",
             "GLUCOSIDES_Sq",
             "TREE.HEIGHT",
             "TREE.HEIGHT_Sq",
             "LEAF.HAIRINESS")

gallTr <- "GALLTYPE"

# Independent term
indep <- paste(c(salixTr, gallTr), collapse = "+")

# Interaction term (without Square)
salixTrNoSq <- c("TREE.VOLUME",
                 "LEAF.THICKNESS",
                 "GLUCOSIDES",
                 "TREE.HEIGHT",
                 "LEAF.HAIRINESS")

gallTrNoSq <- c("GALLTYPE")

combTr <- combn(c(salixTrNoSq, gallTrNoSq), 2)
inter <- paste(combTr[1,], ":", combTr[2,], collapse = "+")

# Build formula
FormulaSalGal <- as.formula(paste("~", indep, "+", inter))

#==========
# Run model
#==========
salixGalGLMTrait <- fitGLM(salixGalSq,
                           formula = FormulaSalGal,
                           family = binomial(link = "logit"))

#============================
# Log-likelihood and Tjur's D
#============================
logLik(salixGalGLMTrait)
tjur(salixGalGLMTrait)

#####################
# Galler - Parasitoid
#####################
# Construct squared traits for numerical traits
galParaSq <- polyTrait(galPara)

#==============
# Build formula
#==============
# Select traits
gallTr <- c("GALLTYPE", 
            "BODYLENGTH.GAL", 
            "BODYLENGTH.GAL_Sq", 
            "PHENOLOGY.GAL", 
            "OVIPOS.STRATEGY",
            "GALL.WALL",
            "GALL.WALL_Sq")

paraTr <- c("P.I",  
            "OVIPOS.LNTH", 
            "OVIPOS.LNTH_Sq", 
            "ATTACK.STAGE",
            "PHENOLOGY.PAR", 
            "BODYLENGTH.PAR", 
            "BODYLENGTH.PAR_Sq", 
            "ENDO.ECTO")

# Independent term
indep <- paste(c(gallTr, paraTr), collapse = "+")

# Interaction term (without Square)
gallTrNoSq <- c("GALLTYPE", 
                "BODYLENGTH.GAL", 
                "PHENOLOGY.GAL", 
                "OVIPOS.STRATEGY",
                "GALL.WALL")

paraTrNoSq <- c("P.I",  
                "OVIPOS.LNTH", 
                "ATTACK.STAGE",
                "PHENOLOGY.PAR", 
                "BODYLENGTH.PAR", 
                "ENDO.ECTO")

combTr <- combn(c(gallTrNoSq, paraTrNoSq), 2)
inter <- paste(combTr[1,], ":", combTr[2,], collapse = "+")

# Build formula
FormulaGalPara <- as.formula(paste("~", indep, "+", inter))

#==========
# Run model
#==========
galParaGLMTrait <- fitGLM(galParaSq,
                           formula = FormulaGalPara,
                           family = binomial(link = "logit"))

#============================
# Log-likelihood and Tjur's D
#============================
logLik(galParaGLMTrait)
tjur(galParaGLMTrait)
```

# `fit4corner`

```{r}
# Construct squared traits for numerical traits
salixGalSq <- polyTrait(salixGal)

#==============
# Build formula
#==============
# Select traits
salixTr <- c("TREE.VOLUME",
             "TREE.VOLUME_Sq",
             "LEAF.THICKNESS",
             "GLUCOSIDES",
             "GLUCOSIDES_Sq",
             "TREE.HEIGHT",
             "TREE.HEIGHT_Sq",
             "LEAF.HAIRINESS")

gallTr <- "GALLTYPE"

# Independent term
indep <- paste(c(salixTr, gallTr), collapse = "+")

# Interaction term (without Square)
salixTrNoSq <- c("TREE.VOLUME",
                 "LEAF.THICKNESS",
                 "GLUCOSIDES",
                 "TREE.HEIGHT",
                 "LEAF.HAIRINESS")

gallTrNoSq <- c("GALLTYPE")

combTr <- expand.grid(salixTrNoSq, gallTrNoSq)
inter <- paste(combTr[,1], ":", combTr[,2], collapse = "+")

# Build formula
FormulaSalGal <- as.formula(paste("~", indep, "+", inter))

#==========
# Run model
#==========
salixGal4cornerTrait <- fitGLM(salixGalSq,
                           formula = FormulaSalGal,
                           family = binomial(link = "logit"))

#============================
# Log-likelihood and Tjur's D
#============================
logLik(salixGal4cornerTrait)
tjur(salixGal4cornerTrait)




#####################
# Galler - Parasitoid
#####################
# Construct squared traits for numerical traits
galParaSq <- polyTrait(galPara)

#==============
# Build formula
#==============
# Select traits
gallTr <- c("GALLTYPE", 
            "BODYLENGTH.GAL", 
            "BODYLENGTH.GAL_Sq", 
            "PHENOLOGY.GAL", 
            "OVIPOS.STRATEGY",
            "GALL.WALL",
            "GALL.WALL_Sq")

paraTr <- c("P.I",  
            "OVIPOS.LNTH", 
            "OVIPOS.LNTH_Sq", 
            "ATTACK.STAGE",
            "PHENOLOGY.PAR", 
            "BODYLENGTH.PAR", 
            "BODYLENGTH.PAR_Sq", 
            "ENDO.ECTO")

# Independent term
indep <- paste(c(gallTr, paraTr), collapse = "+")

# Interaction term (without Square)
gallTrNoSq <- c("GALLTYPE", 
                "BODYLENGTH.GAL", 
                "PHENOLOGY.GAL", 
                "OVIPOS.STRATEGY",
                "GALL.WALL")

paraTrNoSq <- c("P.I",  
                "OVIPOS.LNTH", 
                "ATTACK.STAGE",
                "PHENOLOGY.PAR", 
                "BODYLENGTH.PAR", 
                "ENDO.ECTO")

combTr <- expand.grid(gallTrNoSq, paraTrNoSq)
inter <- paste(combTr[,1], ":", combTr[,2], collapse = "+")

# Build formula
FormulaGalPara <- as.formula(paste("~", indep, "+", inter))

#==========
# Run model
#==========
galParaGLMTrait <- fitGLM(galParaSq,
                           formula = FormulaGalPara,
                           family = binomial(link = "logit"))

#============================
# Log-likelihood and Tjur's D
#============================
logLik(galParaGLMTrait)
tjur(galParaGLMTrait)
```

# `fitIMC` 

This function take a long time to run.

```{r}
# Salix-galler
#salixGalIMC <- fitIMC(salixGal,
#                      d = 1)
```

