---
title: "Salix_Galler_Parasitoids"
author: "Guillaume Blanchet"
date: "14/05/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This vignette was made to be an example of how the package can be applied to real-world data. As such, we focused on the *Salix*-galler-parasitoid data published by Kopelke et al. (2017) for which additional information on phylogeny was added. The full data was used to perform the analysis in Wooton et al. (2021).

# Data

The data is part of the `alien` R package and has already been formatted.

```{r, echo=TRUE, eval=FALSE}
data(salixGal)
data(galPara)
```

# `fitKNN`

## Interaction data only
```{r, warning=FALSE}
# Salix-galler
salixGalKNNInter <- fitKNN(salixGal,
                           distFrom = "jaccard",
                           distTo = "jaccard",
                           nNeig = 3,
                           weight = 0)

ll_fn(salixGal$adjMat,salixGalKNNInter)
tjur_D(salixGal$adjMat, salixGalKNNInter) 
# OK

# Galler-parasitoid
galParaKNNInter <- fitKNN(galPara,
                          distFrom = "jaccard",
                          distTo = "jaccard",
                          nNeig = 3,
                          weight = 0)

ll_fn(galPara$adjMat,galParaKNNInter)
tjur_D(galPara$adjMat, galParaKNNInter) 
# OK
```

## Interaction data with traits
```{r, warning=FALSE}
# Salix-galler
salixGalKNNInterTrait <- fitKNN(salixGal,
                                distFrom = "jaccard",
                                distTo = "jaccard",
                                distTraitFrom = "euclidean",
                                distTraitTo = "euclidean",
                                nNeig = 3,
                                weight = 1)

ll_fn(salixGal$adjMat,salixGalKNNInterTrait)
tjur_D(salixGal$adjMat, salixGalKNNInterTrait) 
# OK

# Galler-parasitoid
galParaKNNInterTrait <- fitKNN(galPara,
                               distFrom = "jaccard",
                               distTo = "jaccard",
                               distTraitFrom = "euclidean",
                               distTraitTo = "euclidean",
                               nNeig = 3,
                               weight = 1)

ll_fn(galPara$adjMat,galParaKNNInterTrait)
tjur_D(galPara$adjMat, galParaKNNInterTrait) 
# OK
```

## Interaction data with phylogeny
```{r, warning=FALSE}
# Salix-galler
salixGalKNNInterPhylo <- fitKNN(salixGal,
                                distFrom = "jaccard",
                                distTo = "jaccard",
                                nNeig = 3,
                                weight = 1,
                                phylo=TRUE)

ll_fn(salixGal$adjMat,salixGalKNNInterPhylo)
tjur_D(salixGal$adjMat, salixGalKNNInterPhylo)
# Not OK

# Galler-parasitoid
galParaKNNInterPhylo <- fitKNN(galPara,
                               distFrom = "jaccard",
                               distTo = "jaccard",
                               nNeig = 3,
                               weight = 1,
                               phylo = TRUE)

ll_fn(galPara$adjMat,galParaKNNInterPhylo)
tjur_D(galPara$adjMat, galParaKNNInterPhylo)
# Not OK
```

# `fitRF`

## Traits
```{r}
# Base of formula
salTraits <- paste(colnames(salixGal$traitFrom), collapse = "+")
galTraits <- paste(colnames(salixGal$traitTo)[1:7], collapse = "+")
parTraits <- paste(colnames(galPara$traitTo), collapse = "+")

# Salix-galler
FormulaSalGal <- as.formula(paste("~", salTraits, "+", galTraits))
salixGalRFTrait <- fitRF(salixGal,
                         formula = FormulaSalGal,
                         ntree = 2000,
                         nodesize = 1)

ll_fn(salixGal$adjMat,salixGalRFTrait)
# OK Small variation is OK because of the stochastic nature of RF

# Galler-parasitoid
FormulaGalPara <- as.formula(paste("~", galTraits, "+", parTraits))
galParaRFTrait <- fitRF(galPara,
                        formula = FormulaGalPara,
                        ntree=2000,
                        nodesize = 1)

ll_fn(galPara$adjMat,galParaRFTrait)

```

## Phylo
This needs to be done after answers from Kate.

# `fitGLM`

```{r}
salixGalTraitSel <- c("TREE.VOLUME",
                      "LEAF.THICKNESS",
                      "GLUCOSIDES",
                      "TREE.HEIGHT",
                      "LEAF.HAIRINESS",
                      "GALLTYPE")

# Salix-galler
FormulaSalGal <- ~ TREE.VOLUME * LEAF.THICKNESS +
                   TREE.VOLUME * GLUCOSIDES +
                   TREE.VOLUME * TREE.HEIGHT +
                   TREE.VOLUME * LEAF.HAIRINESS +
                   TREE.VOLUME * GALLTYPE +
                   LEAF.THICKNESS * GLUCOSIDES +
                   LEAF.THICKNESS * TREE.HEIGHT +
                   LEAF.THICKNESS * LEAF.HAIRINESS +
                   LEAF.THICKNESS * GALLTYPE +
                   GLUCOSIDES * TREE.HEIGHT +
                   GLUCOSIDES * LEAF.HAIRINESS +
                   GLUCOSIDES * GALLTYPE +
                   TREE.HEIGHT * LEAF.HAIRINESS +
                   TREE.HEIGHT * GALLTYPE +
                   LEAF.HAIRINESS * GALLTYPE +
                   I(TREE.VOLUME^2) + 
                   I(GLUCOSIDES^2) + 
                   I(TREE.HEIGHT^2)
  
FormulaSalGal <- ~ TREE.VOLUME + LEAF.THICKNESS +
                   GLUCOSIDES +
                   TREE.HEIGHT +
                   LEAF.HAIRINESS + GALLTYPE +
                   I(TREE.VOLUME^2) + 
                   I(GLUCOSIDES^2) + 
                   I(TREE.HEIGHT^2)
  

FormulaSalGal <- ~ GLUCOSIDES * LEAF.TOUGHN + 
                   GLUCOSIDES * BODYLENGTH.GAL +
                   GLUCOSIDES * DEVELOPMENT.TIME +
                   LEAF.TOUGHN * BODYLENGTH.GAL +
                   LEAF.TOUGHN * DEVELOPMENT.TIME +
                   BODYLENGTH.GAL * DEVELOPMENT.TIME +
                   I(GLUCOSIDES^2) + 
                   I(LEAF.TOUGHN^2) + 
                   I(BODYLENGTH.GAL^2) + 
                   I(DEVELOPMENT.TIME^2)
  
salixGalGLMTrait <- fitGLM(salixGal,
                           formula = FormulaSalGal,
                           family = binomial(link = "logit"))

ll_fn(salixGal$adjMat,salixGalGLMTrait)

```

# `fit4corner`

```{r}
# Salix-galler
FormulaSalGal <- ~ GLUCOSIDES * BODYLENGTH.GAL +
                   GLUCOSIDES * DEVELOPMENT.TIME +
                   LEAF.TOUGHN * BODYLENGTH.GAL +
                   LEAF.TOUGHN * DEVELOPMENT.TIME +
                   I(GLUCOSIDES^2) + 
                   I(LEAF.TOUGHN^2) + 
                   I(BODYLENGTH.GAL^2) + 
                   I(DEVELOPMENT.TIME^2)
  
salixGal4cornerTrait <- fit4corner(salixGal,
                                   formula = FormulaSalGal,
                                   family = binomial(link = "logit"))

ll_fn(salixGal$adjMat,salixGal4cornerTrait)
```

# `fitIMC` 

This function take a long time to run.

```{r}
# Salix-galler
salixGalIMC <- fitIMC(salixGal,
                      d = 1)
```

