---
title: "Salix_Galler_Parasitoids"
author: "Guillaume Blanchet"
date: "14/05/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This vignette was made to be an example of how the package can be applied to real-world data. As such, we focused on the *Salix*-galler-parasitoid data published by Kopelke et al. (2017) for which additional information on phylogeny was added. The full data was used to perform the analysis in Wooton et al. (2021).

# Data

The data is part of the `alien` R package and has already been formatted.

```{r}
library(alien)
library(ape)

data(salixGal)
data(galPara)

#treeGal <- read.nexus("./Tree_with_added_taxa_with_short_codes_from_Mesquite.nex")

#cophGal <- cophenetic(treeGal)
#phyloGal <- as.matrix(salixGal$phyloDistTo)
#sel <- which(rownames(cophGal) %in% rownames(phyloGal))
#cophGalSel <- cophGal[sel,sel]
#cophGalOrd <- cophGalSel[sort(rownames(cophGalSel)),sort(rownames(cophGalSel))]

#salixGal$phyloDistTo <- as.dist(cophGalOrd)
#galPara$phyloDistFrom <- as.dist(cophGalOrd)
```

# `fitKNN`

## Interaction data only
```{r, warning=FALSE}
# Salix-galler
salixGalKNNInter <- fitKNN(salixGal,
                           distFrom = "jaccard",
                           distTo = "jaccard",
                           nNeig = 3,
                           weight = 0)

logLik(salixGalKNNInter, error = 0.001)
tjur(salixGalKNNInter) 

# Galler-parasitoid
galParaKNNInter <- fitKNN(galPara,
                          distFrom = "jaccard",
                          distTo = "jaccard",
                          nNeig = 3,
                          weight = 0)

logLik(galParaKNNInter, error = 0.001)
tjur(galParaKNNInter)
```

## Interaction data with traits
```{r, warning=FALSE}
# Salix-galler
salixGalKNNInterTrait <- fitKNN(salixGal,
                                distFrom = "jaccard",
                                distTo = "jaccard",
                                distTraitFrom = "euclidean",
                                distTraitTo = "euclidean",
                                nNeig = 3,
                                weight = 1)

logLik(salixGalKNNInterTrait, error = 0.001)
tjur(salixGalKNNInterTrait)

# Galler-parasitoid
galParaKNNInterTrait <- fitKNN(galPara,
                               distFrom = "jaccard",
                               distTo = "jaccard",
                               distTraitFrom = "euclidean",
                               distTraitTo = "euclidean",
                               nNeig = 3,
                               weight = 1)

logLik(galParaKNNInterTrait, error = 0.001)
tjur(galParaKNNInterTrait)
```

## Interaction data with phylogeny 
```{r, warning=FALSE}
# Salix-galler
salixGalKNNInterPhylo <- fitKNN(salixGal,
                                distFrom = "jaccard",
                                distTo = "jaccard",
                                nNeig = 3,
                                weight = 1,
                                phylo=TRUE)

logLik(salixGalKNNInterPhylo, error = 0.001)
tjur(salixGalKNNInterPhylo)

# Galler-parasitoid
galParaKNNInterPhylo <- fitKNN(galPara,
                               distFrom = "jaccard",
                               distTo = "jaccard",
                               nNeig = 3,
                               weight = 1,
                               phylo = TRUE)

logLik(galParaKNNInterPhylo, error = 0.001)
tjur(galParaKNNInterPhylo)
```

# `fitRF`

## Traits
```{r}
##############
# Salix-galler
##############

# Base of formula
salTraits <- paste(colnames(salixGal$traitFrom), collapse = "+")
galTraits <- paste(colnames(salixGal$traitTo), collapse = "+")

# Formula
FormulaSalGal <- as.formula(paste("~", salTraits, "+", galTraits))

# Model
salixGalRFTrait <- fitRF(salixGal,
                         formula = FormulaSalGal,
                         ntree = 2000,
                         nodesize = 1)

# Loglikelihood and Tjur's D
logLik(salixGalRFTrait, error = 0.001)
tjur(salixGalRFTrait)

###################
# Galler-parasitoid
###################
# Base of formula
galTraits <- paste(colnames(galPara$traitFrom), collapse = "+")
parTraits <- paste(colnames(galPara$traitTo), collapse = "+")

# Formula
FormulaGalPara <- as.formula(paste("~", galTraits, "+", parTraits))

# Model
galParaRFTrait <- fitRF(galPara,
                        formula = FormulaGalPara,
                        ntree=2000,
                        nodesize = 1)

# Loglikelihood and Tjur's D
logLik(galParaRFTrait, error = 0.001)
tjur(galParaRFTrait)
```

## Phylo
```{r}
library(vegan)
salPhylo <- wcmdscale(salixGal$phyloDistFrom, add = TRUE)
colnames(salPhylo) <- paste0("salPhylo", 1:ncol(salPhylo))

galPhylo <- wcmdscale(salixGal$phyloDistTo, add = TRUE)
colnames(galPhylo) <- paste0("galPhylo", 1:ncol(galPhylo))

paraPhylo <- wcmdscale(galPara$phyloDistTo, add = TRUE)
colnames(paraPhylo) <- paste0("paraPhylo", 1:ncol(paraPhylo))

##############
# Salix-galler
##############
salixGalPhylo <- alienData(adjMat = salixGal$adjMat,
                           traitFrom = as.data.frame(salPhylo),
                           traitTo = as.data.frame(galPhylo))

# Base of formula
salPhyloTraits <- paste(colnames(salixGalPhylo$traitFrom), collapse = "+")
galPhyloTraits <- paste(colnames(salixGalPhylo$traitTo), collapse = "+")

# Formula
FormulaSalGalPhylo <- as.formula(paste("~", salPhyloTraits, "+", galPhyloTraits))

# Model
salixGalRFPhylo <- fitRF(salixGalPhylo,
                         formula = FormulaSalGalPhylo,
                         ntree = 2000,
                         nodesize = 1)

# Loglikelihood and Tjur's D
logLik(salixGalRFPhylo, error = 0.001)
tjur(salixGalRFPhylo)

###################
# Galler-parasitoid
###################
galParaPhylo <- alienData(adjMat = galPara$adjMat,
                           traitFrom = as.data.frame(galPhylo),
                           traitTo = as.data.frame(paraPhylo))

# Base of formula
galPhyloTraits <- paste(colnames(galParaPhylo$traitFrom), collapse = "+")
paraPhyloTraits <- paste(colnames(galParaPhylo$traitTo), collapse = "+")

# Formula
FormulaGalParaPhylo <- as.formula(paste("~", galPhyloTraits, "+", paraPhyloTraits))

# Model
galParaRFPhylo <- fitRF(galParaPhylo,
                        formula = FormulaGalParaPhylo,
                        ntree=2000,
                        nodesize = 1)

# Loglikelihood and Tjur's D
logLik(galParaRFPhylo, error = 0.001)
tjur(galParaRFPhylo)
```

# `fitGLM`

```{r}
################
# Salix - Galler
################
# Construct squared traits for numerical traits
salixGalSq <- polyTrait(salixGal)

#==============
# Build formula
#==============
# Select traits
salixTr <- c("TREE.VOLUME",
             "TREE.VOLUME_Sq",
             "LEAF.THICKNESS",
             "GLUCOSIDES",
             "GLUCOSIDES_Sq",
             "TREE.HEIGHT",
             "TREE.HEIGHT_Sq",
             "LEAF.HAIRINESS")

gallTr <- "GALLTYPE"

# Independent term
indep <- paste(c(salixTr, gallTr), collapse = "+")

# Interaction term (without Square)
salixTrNoSq <- c("TREE.VOLUME",
                 "LEAF.THICKNESS",
                 "GLUCOSIDES",
                 "TREE.HEIGHT",
                 "LEAF.HAIRINESS")

gallTrNoSq <- c("GALLTYPE")

combTr <- combn(c(salixTrNoSq, gallTrNoSq), 2)
inter <- paste(combTr[1,], ":", combTr[2,], collapse = "+")

# Build formula
FormulaSalGal <- as.formula(paste("~", indep, "+", inter))

#==========
# Run model
#==========
salixGalGLMTrait <- fitGLM(salixGalSq,
                           formula = FormulaSalGal,
                           family = binomial(link = "logit"))

#============================
# Log-likelihood and Tjur's D
#============================
logLik(salixGalGLMTrait, error = 0.001)
tjur(salixGalGLMTrait)

#####################
# Galler - Parasitoid
#####################
# Construct squared traits for numerical traits
galParaSq <- polyTrait(galPara)

#==============
# Build formula
#==============
# Select traits
gallTr <- c("GALLTYPE", 
            "BODYLENGTH.GAL", 
            "BODYLENGTH.GAL_Sq", 
            "PHENOLOGY.GAL", 
            "OVIPOS.STRATEGY",
            "GALL.WALL",
            "GALL.WALL_Sq")

paraTr <- c("P.I",  
            "OVIPOS.LNTH", 
            "OVIPOS.LNTH_Sq", 
            "ATTACK.STAGE",
            "PHENOLOGY.PAR", 
            "BODYLENGTH.PAR", 
            "BODYLENGTH.PAR_Sq", 
            "ENDO.ECTO")

# Independent term
indep <- paste(c(gallTr, paraTr), collapse = "+")

# Interaction term (without Square)
gallTrNoSq <- c("GALLTYPE", 
                "BODYLENGTH.GAL", 
                "PHENOLOGY.GAL", 
                "OVIPOS.STRATEGY",
                "GALL.WALL")

paraTrNoSq <- c("P.I",  
                "OVIPOS.LNTH", 
                "ATTACK.STAGE",
                "PHENOLOGY.PAR", 
                "BODYLENGTH.PAR", 
                "ENDO.ECTO")

combTr <- combn(c(gallTrNoSq, paraTrNoSq), 2)
inter <- paste(combTr[1,], ":", combTr[2,], collapse = "+")

# Build formula
FormulaGalPara <- as.formula(paste("~", indep, "+", inter))

#==========
# Run model
#==========
galParaGLMTrait <- fitGLM(galParaSq,
                           formula = FormulaGalPara,
                           family = binomial(link = "logit"))

#============================
# Log-likelihood and Tjur's D
#============================
logLik(galParaGLMTrait, error = 0.001)
tjur(galParaGLMTrait)
```

# `fit4corner`

```{r}
# Construct squared traits for numerical traits
salixGalSq <- polyTrait(salixGal)

#==============
# Build formula
#==============
# Select traits
salixTr <- c("TREE.VOLUME",
             "TREE.VOLUME_Sq",
             "LEAF.THICKNESS",
             "GLUCOSIDES",
             "GLUCOSIDES_Sq",
             "TREE.HEIGHT",
             "TREE.HEIGHT_Sq",
             "LEAF.HAIRINESS")

gallTr <- "GALLTYPE"

# Independent term
indep <- paste(c(salixTr, gallTr), collapse = "+")

# Interaction term (without Square)
salixTrNoSq <- c("TREE.VOLUME",
                 "LEAF.THICKNESS",
                 "GLUCOSIDES",
                 "TREE.HEIGHT",
                 "LEAF.HAIRINESS")

gallTrNoSq <- c("GALLTYPE")

combTr <- expand.grid(salixTrNoSq, gallTrNoSq)
inter <- paste(combTr[,1], ":", combTr[,2], collapse = "+")

# Build formula
FormulaSalGal <- as.formula(paste("~", indep, "+", inter))

#==========
# Run model
#==========
salixGal4cornerTrait <- fitGLM(salixGalSq,
                           formula = FormulaSalGal,
                           family = binomial(link = "logit"))

#============================
# Log-likelihood and Tjur's D
#============================
logLik(salixGal4cornerTrait, error = 0.001)
tjur(salixGal4cornerTrait)


#####################
# Galler - Parasitoid
#####################
# Construct squared traits for numerical traits
galParaSq <- polyTrait(galPara)

#==============
# Build formula
#==============
# Select traits
gallTr <- c("GALLTYPE", 
            "BODYLENGTH.GAL", 
            "BODYLENGTH.GAL_Sq", 
            "PHENOLOGY.GAL", 
            "OVIPOS.STRATEGY",
            "GALL.WALL",
            "GALL.WALL_Sq")

paraTr <- c("P.I",  
            "OVIPOS.LNTH", 
            "OVIPOS.LNTH_Sq", 
            "ATTACK.STAGE",
            "PHENOLOGY.PAR", 
            "BODYLENGTH.PAR", 
            "BODYLENGTH.PAR_Sq", 
            "ENDO.ECTO")

# Independent term
indep <- paste(c(gallTr, paraTr), collapse = "+")

# Interaction term (without Square)
gallTrNoSq <- c("GALLTYPE", 
                "BODYLENGTH.GAL", 
                "PHENOLOGY.GAL", 
                "OVIPOS.STRATEGY",
                "GALL.WALL")

paraTrNoSq <- c("P.I",  
                "OVIPOS.LNTH", 
                "ATTACK.STAGE",
                "PHENOLOGY.PAR", 
                "BODYLENGTH.PAR", 
                "ENDO.ECTO")

combTr <- expand.grid(gallTrNoSq, paraTrNoSq)
inter <- paste(combTr[,1], ":", combTr[,2], collapse = "+")

# Build formula
FormulaGalPara <- as.formula(paste("~", indep, "+", inter))

#==========
# Run model
#==========
galParaGLMTrait <- fitGLM(galParaSq,
                           formula = FormulaGalPara,
                           family = binomial(link = "logit"))

#============================
# Log-likelihood and Tjur's D
#============================
logLik(galParaGLMTrait, error = 0.001)
tjur(galParaGLMTrait)
```

# `fitIMC` 

This function take a long time to run.

```{r}
# Salix-galler
salixGalIMC <- fitIMC(salixGal,
                      d = 1)

saveRDS(salixGalIMC, file = "salixGalIMC.RDS")

logLik(salixGalIMC, error = 0.001) # -137.0956
tjur(salixGalIMC) # 0.526749

# Galler-parasitoids
galParaIMC <- fitIMC(galPara,
                      d = 1)

saveRDS(galParaIMC, file = "galParaIMC2.RDS")
logLik(galParaIMC, error = 0.001) # -690.649; -761.8431
tjur(galParaIMC) # 0.5714072 ; 0.5313928

```

